---
title: 'Feature Selection'
author: "Joaquin Salas"
date: "2020.08.08"
output:
  pdf_document: default
  word_document: default
  
  #data saved in: "featureSelection.RData"
  #save(list = ls(all.names=TRUE), file = "featureSelection.RData")
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This program extract important features useful to distinguish a patient prognosis. Our approach is based on Boruta, a wrapper feature detection method, which feeds from a classifier that delivers a measure of importance, in our case Random Forest.  



```{r  echo=FALSE}




#import libraries

suppressMessages(suppressWarnings(library(Boruta))) #for boruta, feature selection
suppressMessages(library(caret))
suppressMessages(library(caTools))

suppressMessages(library(data.table)) # setnames

suppressMessages(require(dplyr))

suppressMessages(require(dummies)) #dummy variables
suppressMessages(require(ggplot2))
suppressMessages(require("Hmisc"))
suppressMessages(library("installr"))
suppressMessages(library(lubridate, warn.conflict = FALSE, quietly = TRUE))
suppressMessages(library(missForest)) #missForest
suppressMessages(library(matrixStats)) #colSds, standard deviation

#suppressMessages(require(multiROC))
suppressMessages(require(nbpMatching))

suppressMessages(require(plotly))
#suppressMessages(require(PRROC))
#suppressMessages(library(pROC,verbose = TRUE )) #pROC, ROC analysis
suppressMessages(require(plyr))

#suppressMessages(suppressWarnings(library(randomForest))) #for randomForest
suppressMessages(library(readr))

suppressMessages(library(stringi))


suppressMessages(library(tidyr)) #fill








```


## Read the data

To apply this program to your data, you need to redefine the directories where you place 
the data and the code. Also, you need to specify the directory in your local machine where you are placing your figures.  

In the case of Mexico, we have found it convenient to perform this analysis at state level. We use a two letter convention to distinguish among the states, e.g., Queretaro would be QT, Mexico City would be DF and Sinaloa would be SL. Some of these entities are so large that require more more than one file. We suggest to use a pattern to naming files that start with the date and follows my the identity naming convention, e.g., a file name may be '20200620 QT *.csv' 


```{r preparacion, echo=FALSE}


#working directories
code.dir = 'E:/Documents/informs/research/2020.06.25 predicting death/code/'
setwd(code.dir)


filename = paste(code.dir,"selectFeatures.R", sep = "")
source(filename)
filename = paste(code.dir,"acquireData.R", sep = "")
source(filename)

  
#directories with files that  I use across many applications related to COVID
data.dir = 'E://Documents//informs//research//2020.06.25 predicting death//data//'


#directories with files that  I use across many applications related to COVID
figures.dir = 'E://Documents//informs//research//2020.06.25 predicting death//figures//'



#files in the data.dir directory corresponding to the Health Ministery data
prefix = "QT"
files.pattern = paste("20200802 CENSO NOMINAL ",prefix, "*.*", sep = "")

dataset = acquireData(data.dir, files.pattern)


```


## Feature Selection

Using the Boruta technique, we carry out an analysis of the most important characteristics in determining a consistently selected in each iteration. This program is not meant to generate a report by itself but rather, it generates files for further representation. The input includes

* borutaSummaryPublicos_*entity*.RData
The data result of the feature selection process, which may take several hours to compute

* timesSelected_*entity*.png
A graph illustrating the number of times a feature has been selected

* timesSelected_*entity*.csv
A csv file with information about the the number of times a feature has been selected. Meant for further processing 


* importancia_*entity*.csv
Data summarizing the name of the feature deemed important, its mean importance and the standar deviation of the measure of importance over the iterations


* featureSelection_*entity*.RData
A file containing attributes, a variable with the names of the features that were deemed important



```{r seleccion, echo=FALSE}


#run the feature selection n-times
n = 30
#init variables summarizing the n-iterations
cols = dim(dataset$X)[2]
selected = data.frame(matrix(FALSE, nrow = n, ncol=cols))
importance = data.frame(matrix(0, nrow = n, ncol=cols))


setnames(selected,old = names(selected), new= names(dataset$X))

#attempt several times to check for important features
#this process may take several hours
for (i in c(1:n)) {
  print(i)
  
  train = balanceData(dataset$X, dataset$y)
  

  features = selectFeatures(train, prefix)
  selected[i,features$attributes] = TRUE
  importance[i,features$attributes] = features$important$meanImp
}

#number of times a variable has been selected
num.times.selected = colSums(selected,na.rm = TRUE)

#name of the important features
important.attributes = names(selected)[num.times.selected>0]

#mean values 
imp.means = colMeans(importance[,important.attributes], na.rm = TRUE)

#standard deviation of the importance
imp.sd = sapply(importance[,important.attributes], sd, na.rm=TRUE)

#save variables computed so far
filename = paste(data.dir,"borutaSummaryPublicos_",prefix,sep="")
save(num.times.selected, important.attributes, imp.means, imp.sd, selected, importance, file = filename)


#number of times a variable was deemed important sorted in descending order
times.decreasing = order(num.times.selected[names(imp.means)], decreasing = TRUE)
#names of the variables deemed important sorted in descending order
names.decreasing = names(imp.means)[times.decreasing]
print(num.times.selected[names.decreasing])


# Plot the results for importance

filename = paste(figures.dir,"timesSelected_",prefix,".png",sep="")
plot(num.times.selected[names.decreasing], imp.means[names.decreasing], xlab = "veces seleccionado", ylab = "importancia media", col = c(1:length(names.decreasing)), pch = 19, cex = 4)

df = data.frame(veces = num.times.selected[names.decreasing], importancia = imp.means[names.decreasing], nombres = names.decreasing )
filename = paste(data.dir,"timesSelected_",prefix,".csv",sep="")
write.csv(df, filename)
 
dev.off()



#these are features deemded important
imp.features = selected[,important.attributes]





#matrix with unique rows
unique.selected = unique.matrix(selected)

sum = colSums(unique.selected)
attributes.always = names(unique.selected)[sum == dim(unique.selected)[1]]

importance.always = importance[,names(importance) %in% attributes.always]

sapply(importance.always, sd)

sapply(importance.always, mean)



importancia = data.frame(mu = sapply(importance.always, mean), sd = sapply(importance.always, sd) )

importancia <-importancia[order(-importancia$mu),]

filename = paste(data.dir,"importancia_",prefix,".csv",sep="")
write.csv(importancia, filename)



filename = paste(data.dir,"featureSelection_",prefix,".RData",sep="")
attributes = names(importance.always)
save(list = "attributes", file = filename)

#save(list = ls(all.names=TRUE), file = filename)

 

```


