---
title: 'XGBoost Classifier'
author: "Joaquín Salas"
date: "2020.08.08"
output:
  pdf_document: default
  word_document: default
  
  #data saved in: "featureSelection.RData"
  #save(list = ls(all.names=TRUE), file = "featureSelection.RData")
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This program construct a classifier to discriminate between those that live and those that die.
Starting with a dataset, it trains a classifier and evaluate it using ROC and precision-recall curves.
The program divides the dataset in half for training and half for testing. It does so $n$-times. Each time, it analyses the performance. The output files include

* *classifier*\_*entity*\_ROC\_*iteration*.csv
It includes the points in the ROC curve

* *classifier*\_*entity*\_PR\_*iteration*.csv
It includes the points in the precision-recall curve

*classifier* is an acronym where *RF*, *NN*, *SVM*, *XGB* stand for random forest, neural network, support vector machine and extreme gradient boosting. Entity is a two letters word referencing the state, e.g., DF, QT, SL stand for Mexico City, Queretaro and Sinaloa.

*iteration* stands for the number of iteration where the performance was evaluated.


```{r  echo=FALSE}




#import libraries
suppressMessages(library(xgboost))

suppressMessages(library(caret))
suppressMessages(library(caTools))

suppressMessages(library(data.table)) # setnames

suppressMessages(require(dplyr))

suppressMessages(require(dummies)) #dummy variables
suppressMessages(require(ggplot2))
suppressMessages(require("Hmisc"))
suppressMessages(library("installr"))
suppressMessages(library(lubridate, warn.conflict = FALSE, quietly = TRUE))
suppressMessages(library(missForest)) #missForest
suppressMessages(library(matrixStats)) #colSds, standard deviation

suppressMessages(require(multiROC))
suppressMessages(require(nbpMatching))

suppressMessages(require(plotly))
suppressMessages(require(PRROC))
suppressMessages(library(pROC,verbose = TRUE )) #pROC, ROC analysis
suppressMessages(require(plyr))

suppressMessages(suppressWarnings(library(randomForest))) #for randomForest
suppressMessages(library(readr))

suppressMessages(library(stringi))

suppressMessages(library(tidyr)) #fill


suppressMessages(library(e1071))

suppressMessages(library(kernlab)) #svm


#working directories
code.dir = 'E:/Documents/informs/research/2020.06.25 predicting death/code/'
setwd(code.dir)

#directories with files that  I use across many applications related to COVID
data.dir = 'E://Documents//informs//research//2020.06.25 predicting death//data//'



filename = paste(code.dir,"preprocessData.R", sep = "")
source(filename)


#prefix = "QT"
#file.root = "20200802 CENSO NOMINAL "

#prefix = "SL"
#file.root = "20200730  CENSO NOMINAL "

prefix = "DF"
file.root = "20200626 CENSO NOMINAL "

#acquire and preprocess the data to construct a classifier
dataset = preprocessData(prefix, file.root, data.dir)



#seed for the random numbers generator
set.seed(123)


#perfom cross-validation cv.trial times


cv.trial = 30
#accumulate the results in the ROC and precision-recall curve
roc.auc = matrix(0,1, cv.trial)
pr.auc = matrix(0,1, cv.trial)


#divide subset datain half
smp.size = floor(0.5 * nrow(dataset$X))


#cross-validation iterations
for (trial in seq(1,cv.trial)) {
  
  print(trial)
  
  
  
  sample.ind <- createDataPartition(dataset$y, p = 0.5, list = FALSE)
  
  #construct a classifier
  
  train.x = data.matrix(dataset$X[sample.ind,])
  #train.y = as.numeric(dataset$y[sample.ind])-1
  train.y = dataset$y[sample.ind]
  
  test.x = data.matrix(dataset$X[-sample.ind,])
  #test.y = as.numeric(dataset$y[-sample.ind])-1
  test.y = dataset$y[-sample.ind]
  
  
  result = balanceData(train.x, train.y)
  train.x = result$X
  train.y = result$y
  
  
  #construct a classifier
  
  r.eta = 0.1480863
  r.cs = 0.3408778
  r.md = 1
  r.ss = 0.452943
  r.gamma = 0.1494112
  
  r.count = 696
  params=list(eta = r.eta, colsample_bylevel=r.cs,
              subsample = r.ss, max_depth = r.md,
              gamma = r.gamma) 
  
  xgb.classifier <- suppressWarnings(suppressMessages(xgboost(data = train.x, 
                 label =train.y,nrounds = r.count, 
                 numclass = 2, verbose = 0, params = params)))
  
  result = balanceData(test.x, test.y)
  test.x = result$X
  test.y = result$y
  
  
  #predict output. This is what you want to implement in an app
  y.pred = predict(xgb.classifier, test.x)
  
  
  y.pred.n = sigmoid(y.pred*3) #3 is an arbitrary constant. y.pred values range from -0.5 to 1.5

  #performance results
  roc<-roc.curve(scores.class0 = y.pred.n,
                 weights.class0 = test.y, curve= TRUE)
  pr<-pr.curve(scores.class0 =  y.pred.n,
               weights.class0 = test.y, curve= TRUE)
  
  
  #save results for analysis
  roc.auc[ trial] = roc$auc
  pr.auc[ trial] = pr$auc.integral
  
  #save results for each classifier
  filename = sprintf("../data/xgb_%s_roc_%03d.csv", prefix,trial)
  write.csv(roc$curve,filename)
  
  filename = sprintf("../data/xgb_%s_pr_%03d.csv", prefix, trial)
  write.csv(pr$curve,filename)
  
  
  print(roc$auc)
  print(pr$auc.integral)

}






mu = dataset$mu
sigma = dataset$sigma

lista = c("xgb.classifier", "sigma", "mu")
filename = paste(data.dir, "xgbClassifier_",prefix,".RData", sep = "")
save(list = lista, file = filename)

load(file = filename)

```


